---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `{ctsApp}`

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The `{ctsApp}` package provides a graphical user interface (Shiny app) for the `{cts}` (Contraceptives DDI Trial Simulation Platform) package. It offers an intuitive web-based interface to design and simulate drug-drug interactions (DDI) involving contraceptive drugs using physiologically based pharmacokinetic (PBPK) models.

## Installation

You can install the development version of `{ctsApp}` from [GitHub](https://github.com/esqLABS/cstApp) with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("esqLABS/cstApp")
```


## Run

You can launch the Shiny application by running:

```{r, eval = FALSE}
ctsApp::run_app()
```

This will open the application in your default web browser, providing an interactive interface to:

- Import and explore compound models from the OSP model library
- Design DDI simulations between compounds
- Configure dosing protocols and individual characteristics
- Run simulations and analyze results with built-in plots and tables
- Export simulation results for further analysis in PK-Sim
