---
title: "ctsApp User Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ctsApp User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The **ctsApp** package provides a graphical user interface (Shiny app) for the **cts** (Contraceptives DDI Trial Simulation Platform) package. It offers an intuitive web-based interface to design and simulate drug-drug interactions (DDI) involving contraceptive drugs using physiologically based pharmacokinetic (PBPK) models.

### Key Features

- Import and explore compound models from the OSP (Open Systems Pharmacology) model library
- Design DDI simulations between victim and perpetrator compounds
- Configure dosing protocols (oral, IV bolus, IV infusion)
- Define population parameters and individual characteristics
- Run simulations and analyze results with interactive plots
- Export simulation results for further analysis in PK-Sim

## Installation

You can install the development version of ctsApp from GitHub:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("esqLABS/ctsApp")
```

## Launching the Application

To start the application, run:
```{r, eval = FALSE}
ctsApp::run_app()
```

This will open the application in your default web browser.

## User Interface Overview

The ctsApp interface consists of two main areas:

### Left Sidebar

The sidebar contains all configuration panels organized in collapsible accordion sections:

| Panel | Description |
|-------|-------------|
| **Victim** | Configure the victim compound, dosing protocol, and formulation |
| **Population** | Select population type and configure demographics |
| **Perpetrator** | Configure the perpetrator compound, protocol, formulation, and optional EE settings |
| **Simulation Parameters** | Set simulation duration and resolution |

At the bottom of the sidebar, you'll find the **Run Simulation** and **Export Snapshot** buttons.

### Main Content Area

The main area contains three tabs:

| Tab | Description |
|-----|-------------|
| **Experiment Design** | Summary of all configured parameters with demographic visualizations |
| **Results** | Simulation results with PK profiles, DDI analysis, and PK-PD analysis |
| **About** | Information about the application |

## Workflow Guide

### Step 1: Configure Victim Compound

In the **Victim** accordion panel:

1. **Select Compound**: Choose the victim compound from the dropdown. Available options include:
   - Drospirenone (default)
   - Levonorgestrel

2. **Select Protocol**: Choose a predefined dosing protocol or create a new one:
   - Predefined protocols (e.g., "DRSP_3mg 21 days")
   - "Create New Protocol" to define custom dosing

3. **Select Formulation**: Choose a formulation type:
   - Predefined formulations (e.g., "DRSP oral tablet")
   - "Create New Formulation" for custom formulations

### Step 2: Configure Population

In the **Population** accordion panel:

1. **Select Population**: Choose from available population types (e.g., "Healthy Women")

2. **Number of Individuals**: Set the number of virtual individuals (1-100)

3. **Age Range**: Define the age range for the population

4. **Physical Parameters**: Depending on the population, configure either:
   - BMI range, or
   - Height and Weight ranges

### Step 3: Configure Perpetrator Compound

In the **Perpetrator** accordion panel:

1. **Select Compound**: Choose the perpetrator compound:
   - Select from available compounds (e.g., Itraconazole)
   - "Upload Compound" to import a custom compound snapshot (.json)

2. **Select Protocol**: Choose or create a dosing protocol

3. **Select Formulation**: Choose or create a formulation

4. **Ethinylestradiol (EE) Effects** (Optional): Check "Model Ethinylestradiol (EE) effects" to include EE in the simulation. When enabled, additional protocol and formulation options for EE appear.

### Step 4: Set Simulation Parameters

In the **Simulation Parameters** accordion panel:

1. **Duration**: Set the simulation duration and unit (hours, days, weeks)

2. **Resolution**: Set the number of time points per hour

### Step 5: Run Simulation

Once all parameters are configured:

1. The **Run Simulation** button becomes enabled (it remains disabled until all required inputs are valid)

2. Click **Run Simulation** to execute the DDI simulation

3. The simulation creates two scenarios:
   - **Single Simulation**: Victim compound alone
   - **DDI Simulation**: Victim + Perpetrator (+ EE if enabled)

### Step 6: Analyze Results

After the simulation completes, navigate to the **Results** tab to view:

#### Pharmacokinetics Tab

- Time-concentration profiles for all compounds
- Interactive Plotly charts with zoom and pan capabilities
- Comparison between single and DDI simulations

#### PK-DDI Analysis Tab

- DDI ratio calculations
- Comparison of PK parameters (AUC, Cmax, etc.) between scenarios
- Visual representation of drug interaction effects

#### PK-PD Analysis Tab

- Pharmacokinetic-pharmacodynamic analysis (when applicable)

## Advanced Features

### Uploading Custom Compound Snapshots

For the perpetrator compound, you can upload custom compound models:

1. Select "Upload Compound" from the compound dropdown
2. Click "Browse..." to select a JSON snapshot file
3. The snapshot can contain:
   - Compound definitions
   - Formulations
   - Protocols
4. Imported items are automatically added to the available options

### Creating Custom Protocols

When "Create New Protocol" is selected:

| Parameter | Description |
|-----------|-------------|
| **Dose** | Dose amount and unit (mg, g) |
| **Type** | Oral, Intravenous Bolus, or Intravenous |
| **Interval** | Single dose, once/twice/thrice/four times daily |
| **Start Time** | When dosing begins |
| **End Time** | When dosing ends |

For **oral** administration:
- Water Volume/Body Weight: Volume of water co-administered

For **IV** administration:
- Infusion Time: Duration of infusion

### Creating Custom Formulations

When "Create New Formulation" is selected, choose from:

| Type | Description |
|------|-------------|
| **Dissolved** | Immediate release, fully dissolved |
| **Weibull** | Weibull dissolution kinetics |
| **Lint80** | Linear dissolution to 80% |
| **Particle** | Particle-based dissolution (mono/polydisperse) |
| **Table** | Custom dissolution profile via time-fraction table |
| **Zero Order** | Constant release rate |
| **First Order** | Exponential release kinetics |

Each formulation type has specific parameters to configure.

## Exporting Results

### Export DDI Snapshot

Click **Export Snapshot** to download a JSON file containing:

- All compound configurations
- Protocol definitions
- Formulation settings
- Population parameters
- Simulation settings

This snapshot can be:
- Imported into PK-Sim for further analysis
- Shared with collaborators
- Used as a starting point for future simulations

## Troubleshooting

### Buttons are disabled

The **Run Simulation** and **Export Snapshot** buttons are disabled when:

- Required inputs are missing or invalid
- Population parameters are out of valid ranges
- Custom protocol/formulation inputs are incomplete

Check all accordion panels to ensure all required fields are properly configured.
